\pkg{ShinyMICE}: an Evaluation Suite for Multiple Imputation

### 2019-10-30 ResearchReport/article.Rnw
rename folders
Added a simulation folder too
Update research report with quotes from FIMD

@@

- Terminology (MCAR, MAR, MNAR)

@@

The\proglang{R}code and documentation of this project will be open source (available on Github). Since the study does not require the use of unpublished empirical data, I expect that the FETC will grant the label exempt.

@@

@@

@@

"It is often useful to inspect the distributions of original and the imputed data. One way of doing this is to use the function \fct{stripplot} \dots  The differences between the
red points represents our uncertainty about the true (but unknown) values. \dots Under MCAR, univariate distributions of the observed and imputed data are expected to be identical. Under MAR, they can be different, both in location and spread, but their multivariate distribution is assumed to be identical." \cite[p.~12]{mice}. Create scatterplots via \fct{xyplot}.

>>>>

Notation in this paper is copied from \cite{buur18}, and thus based on \cite{rubin87}.

% "Let Y denote the n × p matrix containing the data values on p variables for all n units in the sample. We define the response indicator R as an n × p 0-1 matrix. The elements of Y and R are denoted by y i j and r i j , respectively, where i = 1 , ... , n and j = 1 , ... , p . If y i j is observed, then r i j = 1 , and if y i j is missing, then r i j = 0 ." ... The observed data are collectively denoted by Y o b s . The missing data are collectively denoted as Y m i s , and contain all elements y i j where r i j = 0 . When taken together Y = ( Y o b s , Y m i s ) contain the hypothetically complete data." \cite[par.~2.2]{buur18}

Terminology (MCAR, MAR, MNAR)

% The data are said to be MCAR if Pr ( R = 0 | Y o b s , Y m i s , ?? ) = Pr ( R = 0 | ?? ) (2.1) so the probability of being missing depends only on some parameters ?? , the overall probability of being missing. The data are said to be MAR if Pr ( R = 0 | Y o b s , Y m i s , ?? ) = Pr ( R = 0 | Y o b s , ?? ) (2.2) so the missingness probability may depend on observed information, including any design factors. Finally, the data are MNAR if Pr ( R = 0 | Y o b s , Y m i s , ?? ) (2.3) does not simplify, so here the probability to be missing also depends on unobserved information, including Y m i s itself. \cite[par.~2.2]{buur18}

@@

- missingness mechanims, ignorability

"The practical importance of the distinction between MCAR, MAR and MNAR is that it clarifies the conditions under which we can accurately estimate the scientifically interesting parameters without the need to know $\psi$" \cite[par.~2.2]{buur18}.

@@

The \proglang{R} code and documentation of this project will be open source (available on Github). Since the study does not require the use of unpublished empirical data, I expect that the FETC will grant the label exempt.

@@

See https://stefvanbuuren.name/fimd/sec-sensitivity.html

@@

Create scatterplots via \fct{xyplot} and use lattice functionalities to adapt.

@@

"The mice package contains several graphic functions that can be used to gain insight into the correspondence of the observed and imputed data: \fct{bwplot}, \fct{stripplot}, \fct{densityplot} and \fct{xyplot}." \cite[par.~6.6]{buur18}

see https://onlinelibrary.wiley.com/doi/full/10.1111/1467-9574.00219

"It is often useful to inspect the distributions of original and the imputed data. One way of doing this is to use the function \fct{stripplot} \dots  The differences between the
red points represents our uncertainty about the true (but unknown) values. \dots Under MCAR, univariate distributions of the observed and imputed data are expected to be identical. Under MAR, they can be different, both in location and spread, but their multivariate distribution is assumed to be identical." \cite[p.~12]{mice}.

"Bondarenko and Raghunathan (2016) proposed a more refined diagnostic tool that aims to compare the distributions of observed and imputed data conditional on the missingness probability. The idea is that under MAR the conditional distributions should be similar if the assumed model for creating multiple imputations has a good fit. An example is created as ..."  \cite[par.~6.6]{buur18}

see https://www.ncbi.nlm.nih.gov/pubmed/26952693

% fit <- with(imp, glm(ici(imp) ~ age + bmi + hyp + chl,
%                      family = binomial))
% ps <- rep(rowMeans(sapply(fit$analyses, fitted.values)),
%           imp$m + 1)
% xyplot(imp, bmi ~ ps | as.factor(.imp),
%        xlab = "Probability that record is incomplete",
%        ylab = "BMI", pch = c(1, 19), col = mdc(1:2))
